---
title: "Data Cleaning"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(caret)
library(pROC)
```

## Read in data: 
```{r}
mexico <- read.csv("/Users/sarinakopf/Downloads/Mexico.csv")
dim(mexico)
```

## Restrict to 18-85 Year olds who are citizens of Mexico
```{r}
names(mexico)
data <- filter(mexico, CLASIFICACION_FINAL == 3, EDAD >= 18 & EDAD <=85, NACIONALIDAD == 1)
data <- select(data, c("FECHA_DEF", "NEUMONIA","INTUBADO", "SEXO", "EDAD","INDIGENA", "DIABETES", "EPOC", "ASMA", "INMUSUPR", "HIPERTENSION", "OTRA_COM","CARDIOVASCULAR", "OBESIDAD", "RENAL_CRONICA", "TABAQUISMO"))
```

## Rename columns to english
```{r}
names(data)
colnames(data) <- c("DATE_DEATH", "PNEUMONIA", "INTUBATED", "SEX", "AGE", "INDIGENOUS", "DIABETES", "COPD", "ASTHMA", "IMMUNOSUPR", "HYPERTENSION", "OTHER_DIS", "CARDIOVASCULAR", "OBESITY", "CHRONIC_KIDNEY", "SMOKE")
```


## Make all 97, 98, 99 equal NA
```{r}
##### Pneumonia 
# table(data$PNEUMONIA)
#      1      2     99 
# 169718 863127      9 
data$PNEUMONIA <-  case_when(data$PNEUMONIA == 2 ~ 0, 
                             data$PNEUMONIA == 1 ~ 1, 
                             data$PNEUMONIA == 99 ~ NA_real_)
# table(data$PNEUMONIA)
#      0      1 
# 863127 169718 

##### Intubated 
# table(data$INTUBATED)
#      1      2     97     99 
#  34897 187280 810087    590 
data$INTUBATED <- case_when(data$INTUBATED == 2 ~ 0, 
                            data$INTUBATED == 1 ~ 1, 
                            data$INTUBATED == 97 ~ NA_real_, 
                            data$INTUBATED == 99 ~ NA_real_)
# table(data$INTUBATED)
#      0      1 
# 187280  34897 

##### Sex
# table(data$SEX)
#      1      2 
# 506915 525939 

data$SEX <- case_when(data$SEX == 2 ~ 0, 
                      data$SEX == 1 ~ 1)
# table(data$SEX)
#      0      1 
# 525939 506915 

##### Indigenous
# table(data$INDIGENOUS)
#      1      2     99 
#  10839 988217  33798 
data$INDIGENOUS <- case_when(data$INDIGENOUS == 2 ~ 0, 
                             data$INDIGENOUS == 1 ~ 1, 
                             data$INDIGENOUS == 99 ~ NA_real_)
# table(data$INDIGENOUS)
#      0      1 
# 988217  10839 

##### Diabetes
# table(data$DIABETES)
#      1      2     98 
# 154788 874829   3237 

data$DIABETES <- case_when(data$DIABETES == 1 ~ 1,
                           data$DIABETES == 2 ~ 0, 
                           data$DIABETES == 98 ~ NA_real_)
# table(data$DIABETES)
#      0      1 
# 874829 154788 

##### COPD
# table(data$COPD)
#     1       2      98 
# 13137 1016784    2933 

data$COPD <- case_when(data$COPD == 1 ~ 1, 
                       data$COPD == 2 ~ 0, 
                       data$COPD == 98 ~ NA_real_)
# table(data$COPD)
#       0       1 
# 1016784   13137 

##### Asthma 

# table(data$ASTHMA)
#       1       2      98 
#   25442 1004514    2898 

data$ASTHMA <- case_when(data$ASTHMA == 1 ~ 1, 
                         data$ASTHMA == 2 ~ 0,
                         data$ASTHMA == 98 ~ NA_real_)
# table(data$ASTHMA)
#       0       1 
# 1004514   25442 

##### Immunosuppression 
# table(data$IMMUNOSUPR)
#       1       2      98 
#   10090 1019713    3051 

data$IMMUNOSUPR <- case_when(data$IMMUNOSUPR == 1 ~ 1, 
                             data$IMMUNOSUPR == 2 ~ 0, 
                             data$IMMUNOSUPR == 98 ~ NA_real_)
# table(data$IMMUNOSUPR)
#       0       1 
# 1019713   10090 

##### Hypertension
# table(data$HYPERTENSION)
#      1      2     98 
# 195826 833975   3053 
data$HYPERTENSION <- case_when(data$HYPERTENSION == 1 ~ 1, 
                               data$HYPERTENSION == 2 ~ 0, 
                               data$HYPERTENSION == 98 ~ NA_real_)
# table(data$HYPERTENSION)
#      0      1 
# 833975 195826 

##### Other Diseases 

# table(data$OTHER_DIS)
#       1       2      98 
#   23026 1004758    5070 

data$OTHER_DIS <- case_when(data$OTHER_DIS == 2 ~ 0, 
                            data$OTHER_DIS == 1 ~ 1, 
                            data$OTHER_DIS == 98 ~ NA_real_)

# table(data$OTHER_DIS)
#       0       1 
# 1004758   23026 

##### Cardiovascular

# table(data$CARDIOVASCULAR)
#       1       2      98 
#   18396 1011486    2972 

data$CARDIOVASCULAR <- case_when(data$CARDIOVASCULAR == 1 ~ 1, 
                                 data$CARDIOVASCULAR == 2 ~ 0, 
                                 data$CARDIOVASCULAR == 98 ~ NA_real_)
# table(data$CARDIOVASCULAR)
#       0       1 
# 1011486   18396 

##### Obesity 

# table(data$OBESITY)
#      1      2     98 
# 179465 850450   2939 

data$OBESITY <- case_when(data$OBESITY == 1 ~ 1, 
                          data$OBESITY == 2 ~ 0, 
                          data$OBESITY == 98 ~ NA_real_)
# table(data$OBESITY)
#      0      1 
# 850450 179465 

##### Chronic kidney failure

# table(data$CHRONIC_KIDNEY)
#       1       2      98 
#   18529 1011396    2929 

data$CHRONIC_KIDNEY <- case_when(data$CHRONIC_KIDNEY == 1 ~ 1, 
                                 data$CHRONIC_KIDNEY == 2 ~ 0, 
                                 data$CHRONIC_KIDNEY == 98 ~ NA_real_)
# table(data$CHRONIC_KIDNEY)
#       0       1 
# 1011396   18529 

##### Smoke 
# table(data$SMOKE)
#      1      2     98 
#  78662 951099   3093 

data$SMOKE <- case_when(data$SMOKE == 1 ~ 1, 
                        data$SMOKE == 2 ~ 0, 
                        data$SMOKE == 98 ~ NA_real_)

# table(data$SMOKE)
#      0      1 
# 951099  78662 
```

## Add categorical variable for death 

```{r}
data <- mutate(data, DEATH = ifelse(data$DATE_DEATH == "9999-99-99", 0, 1))
# table(data$DEATH)
#  0      1 
# 935831  97023 
```

## Look at how many people responded NA to multiple explanatory variables 

```{r}
findNAdata <- data[-c(2, 3)]
names(findNAdata)
is.naData <- apply(findNAdata, MARGIN = 2, is.na)
count.na <- apply(is.naData, MARGIN = 1, sum)
table(count.na)
```

```{r}
na.Dataset <- data[count.na > 0, ]
full.dataset <- data[count.na == 0,]
```
The proportions for all the variables in the NA dataset are highly similar to the proportions in the dataset with out any NA responses. We also found that interactions between multiple variables are also unaffected (i.e. Age and Death) by having or removing NAs. We conclude that the NA dataset is largely representative of the complete dataset, and thus by removing all rows with NA values for at least one explanatory variable we are not introducing substantial bias. (See NAexploration.Rmd for full analysis).

## Remove all rows with at least one NA
```{r}
data <- data[count.na==0,]
```

## Remove extra columns
```{r}
data <- data[,-(1:3)]
```

## Split data into testing and training set 
```{r}
## Split data into 50/50 for testing and training
set.seed(9)
inTrain <-  createDataPartition(y = data$DEATH, p = 0.5, list = F)
training <- data[inTrain, ]
testing <- data[-inTrain, ]
dim(training)
dim(testing)
```

## MODEL!!!!
```{r}
full_model <- glm(DEATH~., "binomial", training)
summary(full_model)

reduced <- glm(DEATH~.-CARDIOVASCULAR-INDIGENOUS-ASTHMA-SMOKE, "binomial", training)
summary(reduced)
```

## RESIDUALS!!!~~
```{r}
#might as well include all 4 (doensn't simplify model much, doesn't matter in or out)
full <-  236259
reduced <- 236299

G <- reduced - full
df <- 4

1-pchisq(G, 4)

#make interaction model, see how much that changes (use all interactions)
#after we do it, see how many we can get rid of...
#Focus on AOC curve, cross validation
#In addition to drop in deviance test, do measures of fit 

#statquest package 
#don't look at residual plots 
#Deviance plots kind of already did that 
#Focus on what we have here--multiple explanatory variables (MODEL ASSESMENT )

#Essentially same confusion matrix, g statistic, pick smallest 
#else use 
#AOC, ROC Curves, threshold values, confusion matrix
#You want most area under the curve, pick top left value
#TPP and FPP (true positive/false positive)

#Put people in categories of risk based on age. 
#Set meeting for Monday, maybe also Friday. Free after 10:30.  make another for 



devResids <- residuals(reduced, type="deviance")
pearResids <- residuals(reduced, type="pearson")
trainPlotData <- mutate(training, "pearson" = pearResids, "deviance" = devResids)

# Residual plots for each explanatory variable
ggplot(trainPlotData, aes(x = AGE, y = pearson)) + 
  geom_point(alpha=0.1)+
  labs(title="Residual vs AGE")

ggplot(trainPlotData, aes(x = SEX, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs SEX")

ggplot(trainPlotData, aes(x = CHRONIC_KIDNEY, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs CHRONIC_KIDNEY")  

ggplot(trainPlotData, aes(x = COPD, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs COPD")  

ggplot(trainPlotData, aes(x = IMMUNOSUPR, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs IMMUNOSUPR")

ggplot(trainPlotData, aes(x = OBESITY, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs OBESITY")

ggplot(trainPlotData, aes(x = DIABETES, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs DIABETES")

ggplot(trainPlotData, aes(x = HYPERTENSION, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs HYPERTENSION")

ggplot(trainPlotData, aes(x = OTHER_DIS, y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs OTHER_DIS")

# Residual vs fitted value
ggplot(trainPlotData, aes(x = reduced$fitted.values, y = deviance)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs fitted values")

# Residual vs order
ggplot(trainPlotData, aes(x = 1:nrow(training), y = pearson)) + 
  geom_point(alpha=0.1) +
  labs(title="Residual vs order")

hist(pearResids)
plot(training$AGE, devResids)

median(pearResids[training$SEX == 0])
median(pearResids[training$SEX == 1])

```

## ROC
```{r}
roc(response = training$DEATH, predictor=reduced$fitted.values,ci=F, plot=T)

#confusion counts #predict to die versus #actually die (little 2x2 table)
#they should agree as much as possible 
#probability > 0.5, maybe try 0.8
#calculate area under curve for both
#compare these things for interaction model and new interaction model along with G statistic
#caret can do all things automatically for us 

#odds of success based on a certain age
#DEFINETELY can interpret odds ratio with age (graph, survival curve)
#prob survival, log odds with regression coef. (probability of survival)
#Pearson residual 
#could compare by gender 
#any diease versus no diesase 

# test_prediction <- predict(reduced, testing)
# roc(response = testing$DEATH, predictor=test_prediction,ci=F, plot=T)
```

## Begin! 

```{r}
interaction_model <- glm(DEATH~(.)^2, "binomial", training)
summary(interaction_model)

## Drop - in - deviance test
G <-  full_model$deviance - interaction_model$deviance 
df <- full_model$df.residual - interaction_model$df.residual
1 - pchisq(G, df)

## ROC
roc(response = training$DEATH, predictor=interaction_model$fitted.values,ci=F, plot=T)
#0.8562

0.05 / (interaction_model$df.null - interaction_model$df.residual)

## Reduced interaction
#We used a p-value cutoff of 0.0005 for significance (Adjusted Bonferroni)
interaction_reduced <- glm(data = training, DEATH ~ SEX + AGE + INDIGENOUS + DIABETES + COPD + ASTHMA + IMMUNOSUPR + HYPERTENSION + OTHER_DIS + CARDIOVASCULAR + OBESITY + CHRONIC_KIDNEY + SMOKE + SEX:AGE + SEX:DIABETES + SEX:HYPERTENSION + SEX:CHRONIC_KIDNEY + SEX:SMOKE + AGE:INDIGENOUS + AGE:DIABETES + AGE:COPD + AGE:IMMUNOSUPR + AGE:HYPERTENSION + AGE:OTHER_DIS + AGE:CARDIOVASCULAR + AGE:OBESITY + AGE:CHRONIC_KIDNEY + AGE:SMOKE + DIABETES:OBESITY + IMMUNOSUPR:CHRONIC_KIDNEY + HYPERTENSION:OBESITY + OTHER_DIS:OBESITY , family = "binomial")

summary(interaction_reduced)

#Drop in deviance compared to no interactions 
G <-  full_model$deviance - interaction_reduced$deviance
df <- full_model$df.residual - interaction_reduced$df.residual
1 - pchisq(G, df)


#Drop in deviance compared to ALL interactions
G <-  interaction_reduced$deviance - interaction_model$deviance
df <- interaction_reduced$df.residual - interaction_model$df.residual
1 - pchisq(G, df)

#at least one of interactions we took out is significant 

roc(response = training$DEATH, predictor=interaction_reduced$fitted.values,ci=F, plot=T)
#0.8559

#0.8538 Original Full model with no interaction terms! 
```


## PLOTS PLOTS PLOTS PLOTS PLOTS PLOTS (everybody!)

```{r}

Fits <- mutate(training, fit = interaction_reduced$fitted.values)

ggplot(data = Fits) +
  geom_line(aes(x = AGE, y = fit, color = as.factor(SEX)))

#We are going to have to do plots FOR things. FOR people with obesity...what happens.... (etc)
#As simple as possible
#Choosing some value for everything else, then go from there 

```







## Questions: 

1. Everything significant at 0.05 level except CARDIOVASCULAR. Should we just take this one out, or the four out that didn't have extremely tiny p-values (i.e. CARDIOVASCULAR, ASTHMA, INDIGENOUS, SMOKE), or take out even more variables?
2. We have so many significant p-values. But is this just because we have a huge sample size, so our SEs are small? Would we overfit the model by including all but the four mentioned above?
3. Our residuals tend to be heavily right-skewed. This means that there are a lot more people who died that we didn't expect to die compared to people who survived that we didn't expect to survive. Does this mean that assumptions are violated? How do we fix this?


## DO list: 

1. Model with HELLLAAAAAAAAAAAA interaction terms. Use all terms :( 
2. See if we can take some out 
3. Compare AOC, G statistic, see if improvement is significant. 
4. Find smallest possible model that gives us similar AOC, G-statistic 
5. Once we have our model, do fun stuff! Plot age versus prob survival
  - Odds ratio/Probability changes with different explanatory variables 



